---
title: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Simin Talebpour"
output:
  html_document:
    df_print: paged
---

##Overview

In this project I want to analyze the effect of vitamin C on tooth growth in guinea pigs using ToothGrowth data in the R datasets package. First I'll perform some basic exploratory data analyses, and then use hypothesis tests to compare tooth growth by supplement type and dose level.

## Exploratory data analyses

```{r echo=FALSE, message=FALSE, warning=FALSE,'load required packages'}
# Load required packages 
library(ggplot2)
```


```{r load ToothGrowth Dataset}
data("ToothGrowth")
```


```{r}
str(ToothGrowth)
```


```{r}
summary(ToothGrowth)  
```

```{r}
unique(ToothGrowth$dose)
```

ToothGrowth data includes 60 observations of 3 variables:  
**len**: numeric variable - tooth length   
**supp**: factor variable with 2 levels: "OJ"( Orange Juice) and "VC"(Vitamin C)  
**dose**: numeric valriable	- Dose in milligrams/day  

Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC). The responseis the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs.

```{r, 'box plot for dose'}
ggplot(ToothGrowth, aes(x = factor(dose), y = len)) +
  geom_boxplot(aes(fill = supp))+
  facet_grid(.~supp)+
  ggtitle("Tooth Growth vs. Dose Level for Each Supplement Type")+
  xlab("Dose Level(mg/Day)")+
  ylab("Tooth Length")
      
```

Looking at box plots, I can say by increaing dosage, tooth growth increases for both delivery method(orange juice and Vitamin C).

```{r, 'box plot for supp'}
ggplot(ToothGrowth, aes(x = supp, y = len, fill = factor(dose))) +
  geom_boxplot()+
  facet_grid(.~dose)+
  ggtitle("Tooth Growth vs. Supplement Type for Differnt Dose Levels ")+
  xlab("Supplement Type")+
  ylab("Tooth Length")
```

For lower dosages(0.5, 1) orange juice has been more effective, but for dosage = 2, we can't make a conclusion.

##Hypothesis Tests

I want to see how supplement type and dose level effects the tooth growth. Applying hypothesis tests, I will find my answer.

###Hypothesis One

The supplement type(orange juice or Vitamin C) doesn't effect the tooth growTH.

```{r, 'hypothesis test to compare tooth growth by supp'}
t.test(len ~ supp, ToothGrowth, paired = FALSE, var.equal = FALSE)

```

Since the p-value(0.06) is greater than my significant value(0.05), I fail to reject the null hypothesis.

###Hypothesis Two

The dose levels '1' and '2' have the same effect on the tooth growTH.

```{r, 'hypothesis test to compare tooth growth by dose 1 and 2'}
ToothGrowth12 <- subset(ToothGrowth, dose %in% c(1,2))
t.test(len ~ dose, ToothGrowth12, paired = FALSE, var.equal = FALSE)

```

P-value(1.906e-05) is smaler than my significance value(0.05), so I reject the null hypothesis. Tooth growth is not the same for dose = 1 and dose = 2.

###Hypothesis Three

The dose levels '1' and '0.5' have the same effect on the tooth growTH.

```{r, 'hypothesis test to compare tooth growth by dose 0.5 and 1'}
ToothGrowth105 <- subset(ToothGrowth, dose %in% c(1,0.5))
t.test(len ~ dose, ToothGrowth105, paired = FALSE, var.equal = FALSE)

```

P-value(1.268e-07) is smaler than my significance value(0.05), so I reject the null hypothesis. Tooth growth is not the same for dose = 1 and dose = 0.5.

###Hypothesis Four

The dose levels '2' and '0.5' have the same effect on the tooth growth.

```{r, 'hypothesis test to compare tooth growth by dose 0.5 and 2'}
ToothGrowth205 <- subset(ToothGrowth, dose %in% c(2,0.5))
t.test(len ~ dose, ToothGrowth205, paired = FALSE, var.equal = FALSE)

```

P-value(4.398e-14) is smaler than my significance value(0.05), so I reject the null hypothesis. Tooth growth is not the same for dose = 2 and dose = 0.5.

##Conclusion

I investigated ToothGrowth dataset to see if there is any differnce in guinea pig's tooth growth based on the type and dose level of vitamin C supplement they recieved. To naswer this question I performed four independent two-sample t-tests assuming the followings:  
- The two samples in each test are independent.  
- The two samples in each test follow normal distributions.  
- The two samples in each test have differnt variance.  

These are my findings:  
The supplement type doesn't effect the tooth growth, however the extra dose increases the tooth growth.



